---
title: "Documentation of the extension of the fisheries composition dataset"
author: "Segovia D L Revilla, Lucia"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)



```

## SUMMARY

**insert image**

Flowchart:

1)  Identifying the data:
    QC1: Evaluate the quality of the dataset and the documentation
    accompanying the dataset.
    
2) Importing the data into R:
    This step is going to depend on the original (raw) format
    QC2: We will ensure that all the data has been imported and done correctly. E.g., check that total number of observations and variables are correctly imported.
    
3) Tidying the data:
    This means we will convert from the original (raw) format to a standardised (harmonised) format, by performing all or some of the following operations:
        a. Variable names: Components will be name using Tagnames (upper case) and the standard unit (lower case) (e.g., PROTg).
        b. Diving combined variables into two (or more) columns. E.g., CARTEB or [CARTEBEQ] into CARTEB, CARTBEQ. 
        c. *Removing brackets* or other special characters ("[]", "()", "\*"). E.g., extracting the numeric value from inside the bracket  from "low quality"/alternative methods. --> How can we record/ account for it within the data?
        d. *Removing summary stat from the table* --> How can we keep/use that data?
        e. Food name/ description standardisation: FAO ICP codes will be assigned to fishery products.
        f. Food groups variable (row to column)
    QC3: Check that observations and variables are still the same as the original dataset. 

4) Data transformation: 
    Some variables will be recalculated, for example, energy will be recalculated. Summary statistics will be calculated for each FAO ICP codes. 
        a. Changing characters into numeric. For example, trace ("tr") or below detection limit will be converted into zero (0). 
        b. Changing variable units. For example from mg to g, according to the standard reporting units (FAO/INFOODS, 2013).
        c. Re-calculating variables: For example, energy needs to be calculated from other components (macronutrients).

5) Visualising the data: missing values and outliers will be plotted and identified. 

6)  Exporting data and metadata

## Background

**complete this w/ info from the contract and report**

### 1) Preparation steps: Identifying the data

#### 1. 1) Reviewing previous work by FAO team on the current version of the Fisheries Global NCT.

a)  Identifying datasets and data availability:

Of the 13 FCTs included in the Global Table of Nutrient Values, for the Fisheries Global NCT, 2022 (FISHERIES-GlobalNCT_ForSharing_Feb2022.xls), uPulses FCTs was
excluded because only reported pulses. Then, three FCTs were not
available and, we requested access to the FAO Team, this was covered by the
data sharing agreement.

Then, we reviewed the availability of ten pre-selected food components
in the 12 FCTs, and we identified some data gaps, hence we identified other
possible sources of nutrient values (e.g. The Norwegian FCDB, 2021) to
be reviewed and approved by the FAO team.

b)  Checking quality assessment of the NVs:

Because most of the FCTs/FCDBs included were previously reviewed by the
FAO team, there was no need of checking the quality of the data.

#### 1. 2) Identifying new potential sources of NVs 

We reviewed some FCTs that could provide information on fish and fishery products for the pre-defined nutrients of interest. Some potential candidates were FCTs of countries or region with potentially high consumption of fish: Ciqual FCDB, 2019, Norway FCDB, 2021, ASEAN FCDB, 2014. 

a)  Assessing quality and availability of data: We used the FAO/INFOODS
    framework (screening questions) but adapted for fish to identify
    potential source of nutrient values. This led to the inclusion of
    the Norwegian FCDB, 2021. Because, it was the only new FCDB included
    which was not previously reviewed by the FAO team, we checked the
    original values of the Norwegian FCDB, 2021 (NO21) in depth, which
    is discussed in the **Section 2.1** of the protocol of the NO21
    (NO21_docu.Rmd).

In table 1, we can find information on the final list of FCDBs included.

#### Table 1. 

| BiblioID | Reference | Link to access original data | Access | Format | Reviewer |
|------------|------------|------------|------------|------------|------------|
| US19     | US Department of Agriculture (USDA), Agricultural Research Service, Nutrient Data Laboratory. USDA National Nutrient                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | Database for Standard Reference, Legacy. Version Current: April 2019.                       | <http://www.ars.usda.gov/nutrientdata>                                                       | Publicly available | Access    | FAO|
| AU19     | Food Standards Australia New Zealand (FSANZ). 2019. The Australian Food Composition Database, release 1 [online].                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              | Canberra.                                                                                   | <http://www.foodstandards.gov.au>                                                            | Internal           | excel     | FAO|
| NZ18     | New Zealand Food Composition Database. 2019. New Zealand FOODfilesTM 2018 Manual. The New Zealand Institute for                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                | Plant and Food Research Limited and Ministry of Health.                                     | www.foodcomposition.co.nz/foodfiles                                                          | Publicly available | excel     | FAO|
| UK21     | Food Standards Agency. 2021. McCance and Widdowson's The Composition of Foods Integrated Dataset (CoFID) 2021.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 | London, Institute of Food Research, Public Health England.                                  | <https://www.gov.uk/government/publications/composition-of-foods-integrated-dataset-cofid>). | publicly available | excel     | FAO|
| DK19     | Frida, DTU Foods public food database, version 4, 2019, National Food Institute, Technical University of Denmark.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              | <https://frida.fooddata.dk>                                                                 | Publicly available                                                                           | excel              | FAO       |
| WA19     | Vincent, A., Grande, F., Compaoré, E., Amponsah Annor, G., Addy, P.A., Aburime, L.C., Ahmed, D., Bih Loh, A.M., Dahdouh Cabia, S., Deflache, N., Dembélé, F.M., Dieudonné, B., Edwige, O.B., Ene-Obong, H.N., Fanou Fogny, N., Ferreira, M., Omaghomi Jemide, J., Kouebou, P.C., Muller, C., Nájera Espinosa, S., Ouattara, F., Rittenschober, D., Schönfeldt, H., Stadlmayr, B., van Deventer, M., Razikou Yiagnigni, A. & Charrondière, U.R. 2020. FAO/INFOODS Food Composition Table for Western Africa (2019) User Guide & Condensed Food Composition Table / Table de composition des aliments FAO/INFOODS pour l'Afrique de l'Ouest (2019) Guide d'utilisation & table de composition des aliments condensée. Rome, FAO. |                                                                                             | Publicly avaliable                                                                           | excel              | FAO       |
| KE18     | FAO & Government of Kenya. 2018. Kenya Food Composition Tables [online]. Nairobi.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              | <http://www.kilimo.go.ke/wp-content/uploads/2018/10/KENYA-FOOD-COMPOSITION-TABLES-2018.pdf> | Publicly available                                                                           | excel              | FAO       |
| IN17     | Longvah, T., Ananthan, R., Bhaskarachary, K. & Venkaiah, K. 2017. Indian Food Composition Tables. Hyderabad, India, National Institute of Nutrition.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           | <http://www.ifct2017.com/frame.php?page=home>).                                             | Internal                                                                                     | excel              | FAO       |
| JA15     | MEXT. (2015).The Standard Tables of Food Composition in Japan 2015 (SeventhRevised Edition). Official Gazette Co-operation of Japan. Japan                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     | <http://www.mext.go.jp/a_menu/syokuhinseibun/1365451.htm/>                                  | Publicly available                                                                           | excel              | FAO       |
| BA13     | Shaheen N., Rahim A.T.M.A., Mohiduzzaman Md., Banu C.P., Bari Md. L., Tukun A.B., Mannan M.A., Bhattacharjee L., Stadlmayr B. 2013. Food Composition Table for Bangladesh. Institute of Nutrition and Food Science, Centre for Advanced Research in Sciences, University of Dhaka.                                                                                                                                                                                                                                                                                                                                                                                                                                             |                                                                                             | Publicly available                                                                           | excel              | FAO       |
| BR11     | Nucleo de Estudos e pesquisas em Alimentacao (NEPA). Brazilian Food Composition Table (TACO), 4th ed., 2011.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   | <https://www.nepa.unicamp.br/taco/tabela.php?ativo=tabela>                                  | Internal                                                                                     | excel              | FAO       |
| UF16     | FAO. 2016. FAO/INFOODS Global Food Composition Database for Fish and Shellfish. Version 1.0 -- uFiSh1.0. Rome.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 | <http://www.fao.org/infoods/infoods/tables-and-databases/faoinfoods-databases/en/>).        | Publicly available                                                                           | excel              | FAO       |
| UP17     | FAO. 2017. FAO/INFOODS Global Food Composition Database for Pulses. Version 1.0 -- uPulses 1.0. Rome.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          | <http://www.fao.org/infoods/infoods/tables-and-databases/faoinfoods-databases/en/>).        | Publicly available                                                                           | excel              | FAO       |
| NO21     | Norwegian Food Composition Database 2021. Norwegian Food Safety Authority.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     | <https://www.matportalen.no/verktoy/the_norwegian_food_composition_table/>                  | Publicly available                                                                           | excel              | UoN       |

Note: UP17 was not process because we were only interested in fish.

## Harmonisation of the food composition tables and databases (method)

The harmonisation required two main steps: importing (loading) databases
into R and cleaning and standardising the databases.


### 2) Importing the data:

Datasets were found in a variety of formats that needed to be
imported and cleaned into a tabular format. E.g. one row per food entry
and one column per variable (food entry attribute).

The steps and scripts for importing each FCDB are stored together with the original FCDB (raw files) in individual folders named with the BiblioID (See table 1) which is stored in the variable `source_fct`.

### 3) Tidying the data

In order to merge all data from various FCDBs we needed to harmonise the
information in each dataset. The steps that were taken are
explained below and annotated scripts for each FCDBs can be found in
this repository in each corresponding folder.

#### Renaming variable names

Each food composition database (FCDB) has its own variable names,
including the nutrient values. Some FCDB included information related
to the Tagnames
[<https://www.fao.org/infoods/infoods/standards-guidelines/food-component-identifiers-tagnames/en/>]
while other did not. In order to merge all data from various FCDBs we
need to harmonise the names of all the variables and particularly the
food components of interest. To do so, we evaluated and renamed it
according to the Tagnames, while for other variables, we standardised
them by, for instance, recalculating them (e.g., Edible portion).

Other common variables (e.g., food id, food name or food description)
were renamed to a common variable name, for instance, fdc_id, food_desc,
following the variable names used in the Fisheries Global NCT file.

###### The identification of food components

Information on the food components and their description should be
sought in each FCT included. We are reviewing the FAO/ INFOODS food
components identifiers  (Tagnames) used in the previous version for
generating the same structure and format in our output table. Some minor
changes in the variables names were introduced to be compliant with R conventions.
I.e., removing spaces in variables names, changing µg to mcg,
standardising the "standardised" name formatting, from using underscores and/or parenthesis to using only underscores. Also, changing
dashes (-) to underscores (-). We also assumed that all the standardised
variables are recalculated. Note that normally Tagnames uses the dash to denote that the method for obtaining that (component) value is unknown. This is important for the quality
assessment of both, the (fish) items and the overall FCDB quality.

*Table 2* provide a list of all the variables including components used  in the previous version of the Fisheries Global NCT. 


**Table 2.**: List of variables names (Tagnames), description and reference for the Tagnames and description.

|Variable name| Description| Reference|
|-|-|-|
|fdc_id | food identifier as per original source (FCT) |-|
|food_desc | food name and description as per original source (FCT) |-|
|food_group| food group as per original source (FCT)|-|
|scientific_name | Scientific name of the food |-|
|source_fct | Id. of the FCT (BiblioID), see Table 1 |-|
|nutrient_data_source | references or data source for the nutrient values |-|
|Edible_factor_in_FCT | edible portion reported as fraction edible, imputed from the original FCT, or calculated from refuse in the FCT |-|
|Edible_desc | information provided on the edible portion |-|
|specific_gravity | density reported in the original units as in the origianl FCT|-|
|SOPg | Sum of Proximate in g per 100g EP as reported in the original FCT |-|
|ASHg | Ashes in g per 100g of EP|-|
|ENERCkJ | Energy in kJ/ 100g of EP as reported in the original FCT |-|
|ENERCkcal | Energy in kcal/ 100g EP as reported in the original FCT |-|
|WATERg | Water/ moisture content in g per 100g of EP|-|
|PROCNTg | protein in g per 100g of EP, as reported in the original FCT and assumed to be calculated from NT content| FAO/INFOODS, 2012|
|NTg | Nitrogen, total|FAO/INFOODS, 2012|
|XN | Nitrogen conversion factor|FAO/INFOODS, 2012|
|FATg | total fat content in g per 100g of EP. Sum of triglycerides, phospholipids, sterols and related compounds obtained by a mixed solvent extraction |FAO/INFOODS, 2012|
|FAT_g | fat content unknown method of calculation in g per 100g of EP|-|
|FATCEg | fat content measured using continuous extraction (e.g., Soxhlet method) expressed in g per 100g of EP|FAO/INFOODS, 2012|
|CHOAVLDFg | available carbohydrates calculated by difference (E.g., 100 - (weight in grams of water + protein + fat + ash + fibre + alcohol in 100 g of food)) imputed from the original source |FAO/INFOODS, 2012|
|CHOAVLg | available carbohydrates calculated by weight in g per 100g of EP |FAO/INFOODS, 2012| |CHOCDFg | total carbohydrates calculated by difference (E.g., 100 - (weight in grams of water + protein + fat + ash + alcohol in 100 g of food)), as reported in the FCT |FAO/INFOODS, 2012|
|CHOCSMg  | carbohydrates, total; calculated by summation. This values is the sum of the sugars, starches, oligosaccharide, and dietary fibre in g per 100g of EP|FAO/INFOODS, 2012|
|CHOAVLMg | carbohydrates, available in monosaccharide equivalent. This value includes the free sugars plus dextrins, starch, and glycogen in g per 100g of EP |FAO/INFOODS, 2012|
|FIBTGg | total dietary fibre by AOAC Prosky method expressed in g per 100g of EP |FAO/INFOODS, 2012|
|NSPg | non-starch polysaccharide, (Englyst fibre) expressed in g per 100g of EP|FAO/INFOODS, 2012|
|FIBCg | fibre, crude in g per 100g of EP |FAO/INFOODS, 2012|
|ALCg | alcohol in g per 100g |-|
|ALCg_100mL | alcohol expressed in g per 100mL |-|
|SUGARg | total sugar (the sum of free mono- and disaccharides) in g per 100g of EP |-|
|FASATg | fatty acids, total saturated in g per 100g |FAO/INFOODS, 2012|
|FAMSg  | fatty acids, total monounsaturated in g per 100g |FAO/INFOODS, 2012|
|FAPUg  | fatty acids, total polyunsaturated in g per 100g of EP |FAO/INFOODS, 2012|
|FATRNg | fatty acids, total trans in g per 100g EP |FAO/INFOODS, 2012|
|F22D6N3g | docosahexaenoic acid (DHA) (22:6 n-3) in g/100g of EP |-|
|F20D5N3g | eicosapentaenoic acid (EPA) (20:5 n-3) in g/100g of EP |-|
|CHOLEmg | cholesterol determined by enzymatic or chromatographic method in mg per 100g of EP|-|
|CHOL_mg | cholesterol by unknown method of determination in mg per 100g of EP|-|
|RETOLmcg | retinol in mcg per 100g of EP |-|
|VITAmcg | Vitamin A (Retinol Eq. (RE) = mcg retinol + 1/6 mcg ß- carotene + 1/12 mcg other pro-vitamin A carotenoids) in mcg per 100g of EP |FAO/INFOODS, 2012|
|VITA_RAEmcg | Vitamin A (Retinol Activity Eq. (RAE) = mcg retinol + 1/12 mcg ß- carotene + 1/24 mcg other provitamin A carotenoids) in mcg per 100g of EP |FAO/INFOODS, 2012|
|CARTBEQmcg | beta-carotene equivalents, is the sum of the beta-carotene + 1/2 quantity of other carotenoids with vitamin A activity, expressed in mcg per 100g of EP |FAO/INFOODS, 2012|
|CARTAmcg | alpha-carotene in mcg per 100g of EP |FAO/INFOODS, 2012|
|CARTBmcg | beta-carotene in mcg per 100g of EP |FAO/INFOODS, 2012|
|CRYPXBmcg| beta-cryptoxanthin in mcg per 100g of EP |FAO/INFOODS, 2012|
|VITEmg | vitamin E calculated by summation of the vitamin E activities of the active tocopherols and tocotrienols expressed as alpha-tocopherol equivalent in mg per 100g of EP (alpha-tocophero eq. = alpha-tocopherol + 0.4 beta-tocopherol + 0.1 gamma-tocopherol+ 0.01 delta-tocopherol+ 0.3 alpha-tocotrienol + 0.05 beta-tocotrienol + 0.01 gamma-tocotrienol) | FAO/INFOODS, 2012 |
|TOCPHAmg| alpha-tocopherol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|TOCPHBmg | beta-tocopherol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|TOCPHGmg | gamma-tocopherol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|TOCPHDmg | delta-tocopherol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|TOCTRAmg | alpha-tocotrienol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|TOCTRBmg | beta-tocotrienol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|TOCTRGmg | gamma-tocotrienol in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|THIAmg | thiamin, vitamin B1 analysed and expressed as thiamin in mg per 100g of EP | FAO/INFOODS, 2013b|
|THIAHCLmg | thiamin hydrochloride, vitamin B1 analysed and expressed as thiamin hydrochloride in mg per 100g of EP | FAO/INFOODS, 2013b|
|RIBFmg | Riboflavin (Vitamin B2) in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|VITB6Amg | Vitamin B6 determined by analysis in mg/100g of EP | FAO/INFOODS, 2012 |
|VITB6Cmg | Vitamin B6 determined by calculation (sum of pyridoxal, pyridoxamine and pyridoxine) in mg per 100g EP | FAO/INFOODS, 2012|
|VITB6_mg | Vitamin B6 by unknown method in mg/100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|FOLDFEmcg | dietary folate equivalents calculated as food folate + 1.7 x synthetic folic acid and expressed in mcg per 100g of EP | FAO/INFOODS, 2012|
|FOLmcg | folate, total (food folate + synthetic folic) in mcg per 100g EP. Includes both conjugated and free folate (determined by microbiological assay) | FAO/INFOODS, 2012|
|FOLACmcg | folic acid. Synthetic folic acid used in fortification and expressed in mcg per 100g EP | FAO/INFOODS, 2012|
|FOLFDmcg | folate food, naturally occurring food folate (determined by microbiological assay) in mcg per 100g EP | FAO/INFOODS, 2012|
|FOLSUMmcg | folate, sum vitamers in mcg per 100g EP. It includes mostly tetrahydrofolate, 5-methyltetrahydrofolate, 5-formyltetrahydrofolate, 10-formylfolic acid, 10-formyldihyrdofolate and folic acid (determined by HPLC)
|FOL_mcg | folate total, unknown method | Fisheries Global NCT, 2022|
|NIAEQmg | niacin equivalents, total. Preformed niacin plus niacin equivalents from tryptophan (TRP) in mg per 100g EP  | FAO/INFOODS, 2012|
|NIAmg | Niacin, prefrormed in mg per 100g EP  | FAO/INFOODS, 2012|
|NIATRPmg | niacin equivalents, from tryptophan. 1/60 x tryptophan (TRP) expressed in mg per 100g EP  | FAO/INFOODS, 2012|
|TRPmg = tryptophan in mg per 100g of EP (includes only L-tryptophan)| Tagnames, 2007 (accessed on 2022-06-08)|
|VITB12mcg | Vitamin B12 (cobalamin, including all the active forms in foods) expressed in mcg per 100g EP | Tagnames, 2007 (accessed on 2022-06-08)|
|VITCmg | vitamin C (L-ascorbic acid plus Ldehydro-ascorbic acid). Usually analyzed by HPLC and expressed in mg per 100g of EP | FAO/INFOODS, 2012
|ASCLmg | L-ascorbic acid in mg per 100g of EP. Titrimetry can normally analyze L-ascorbic acid only | FAO/INFOODS, 2012|
|VITDEQmcg | vitamin D calculated as the sum of Vitamin D3 + D2 + 5 x 25-hydroxycholecalciferol and expressed in mcg per 100g of EP | FAO/INFOODS, 2012|
|VITDmcg | vitamin D calculated by summation of ergocalciferol and cholecalciferol and expressed in mcg per 100g of EP. This definition is mostly used | FAO/INFOODS, 2012|
|CHOCALmcg | cholecalciferol (D3) in mcg per 100g of EP | FAO/INFOODS, 2012|
|ERGCALmcg | ergocalciferol (D2)  in mcg per 100g of EP | FAO/INFOODS, 2012|
|CHOCALOHmcg | 25-hydroxycholecalciferol in mcg per 100g of EP | FAO/INFOODS, 2012|
|ERGCALOHmcg | 25-hydroxyergocalciferol | Own definition based on CHOCALOHmcg|
|CAmg | calcium in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|MGmg | magnesium in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|MNmg | manganese in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08) |
|Pmg | phosphorus in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|FEmg | iron in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|NAmg | sodium in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|Kmg | potassium in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|CUmg | copper in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|ZNmg | Zinc in mg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08) |
|SEmcg | selenium in mcg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|
|IDmcg | iodine in mcg per 100g of EP | Tagnames, 2007 (accessed on 2022-06-08)|

We extracted information on all the nutrients available in each FCDB,
however more attention have been given to the following list of
nutrients: DHA, EPA, Vitamin B6, Vitamin B12, Vitamin D, Niacin, Iodine, Selenium, Copper, and Sodium.   

##### Components that we need to collect information for:

   1. Iodine - excluded
   2. Se - included
   3. Cu - included
   4. Na? - excluded 
   5. Niacin (Vit. B3) - included, if we are calculating NIATRP, we need TRP too. 
   6. Vit. B6 - included, VITB6A = from analysis, VITB6C = calculated, and VITB6-.
   7. Vit. B12 - included
   8. Vit. D - excluded (including D3, D2 and hidroxi forms).
   9. PUFA - DHA (22:6 (n−3)) - included 
   10. PUFA - EPA (20:5 (n-3)) - included 

#### Other tidying operations:

		**a. Diving combined variables into two (or more) columns.** For example, when two components (Tagnames) were reported in the same column and identified using brackets ([]) around the component values. (CARTEB or [CARTEBEQ]) were separated into two independent columns and bracket removed (CARTEB, CARTBEQ). 
		**b. Removing brackets or other special characters ("[]", "()", "*")**. Similarly, special character were used to denote "low quality values" and/or alternative (determination) methods. For those cases, the numeric value was extracted and the special character removed. For future development, a way of keeping that information should be found (e.g, how can we record and account for those low quality values within the dataset?, can we generate a score with the number of "low quality" values per food item?).
		**c. Removing summary statistics from the table.** Because summary statistics will not be used were removed from the datasets. For future development, a way to account for variability (e.g., using the summary stats) could be studied (How can we keep/use that data?, could we use them to account for uncertainty in the final results?). 
		**d. Food name/ description standardisation.** The ASFIS list was used for identification of the fish and fishery products. The ISSCAAP and the 3-alpha codes were assigned to fishery products for identification and name standardisation. The ISSCAAP codes were also used for assigning the FAO ICP code. The FAO ICP code were extracted from previous work reported in Fisheries Global NCT, 2022, except for NO21 which is reported in (NO21_docu.Rmd)
		**e. Food groups variable.** Some food composition tables reported food groups that were placed as the first row of each category. Those rows were extracted and allocated as a new attribute of each food/fish entry (i.e., from row to new column (`food_group`)).


### 4) Data transformation

Some variables needed more than just renaming to allow comparison.

#### 4.1) Changing characters into numeric

To perform mathematical operations, characters needed to be converted into numeric operator. For instance, values that were reported to be trace ("tr") or below the detection limit ("<LOD") were converted to zero (0).

#### 4.2)  Changing variable units

To standardise and merge the different FCT, components needs to be reported in the same units. For example, some nutrients needed to be convert from mg to g, or from percentage (100%) to a fraction (1) in edible portion. For all the unit conversion we followed the FAO/INFOODS Guidelines for Converting Units, Denominators and Expression (FAO/INFOODS, 2012b), and the suggested standard reporting units.

Eg. Converting alcohol from weight in volumen (w/v) to weight in mass. 

**Eq.1** ALC (g/100mL) / density (g/mL) = ALC (g/100 EP)

Eg. amino acids (AA) reported per g in 100g of PROT to mg in 100g of EP:

 **Eq.2.1** AA mg/100g EP = AA mg/g prot \* prot g/100g EP /100

 **Eq.2.2** AA mg/100g EP = AA g/ 100g prot \* prot g/100g EP /100 * 1000/100 
 **Eq.2.3** AA mg/100g EP = AA g/100g prot  \* prot g/100g EP /10

#### 4.3)  Re-calculating variables

For example Energy needs to be calculated from other components (macronutrients) or edible portion from refuse.
 
 **Eq.3.1** Edible = 100-Refuse(%)/100
 **Eq.3.2** Edible = 1-Refuse
 
#### 4.4) Combining Tagnames

Some components could be reported using more than one Tagnames (as described in the renaming section). Of the fish-relevant nutrients selected for review some could be expressed using different Tagnames, depending on: the fraction
analysed (e.g., Vitamin D3, Vitmin D2), or the method of analysis (e.g., VITB6A, VITB6C). When possible, they were standardised by combining values from similar Tagnames such as Vitamin B6 into a `_standardised` variable.


##### Generating functions for data transformation

**Special values**

We generate a function (`no_brackets_tr`) to be called to change trace and <LOD values to zero and remove brackets (see functions.R).

Next step to disaggregate the function into two functions and to generate a new column for reporting the changes (see Section: Next steps). 

-   Tr, LOD == 0
-   Brackets


# Updating the Global Fish NCT:

Once the FCTs/FCDBs were loaded and harmonised, we combined them into
a single data file, filter the fish entries and merged with the
information already provided by the FAO team regarding each fish entry
ISSCAAP, ICS FAOSTAT and alpha-three codes.

For this particular case the steps 4.4 and 4.5 were performed after merging all the datasets together.

All the steps are done and documented in the following script. 


#Results

### 2) Importing FCTs

#### Renaming variables

Checking food component names and renaming them with Tagnames

Some FCDBs reported the Tagnames which could be found in different
formats.

-   In the Australia FCDB, 2019, it was expressed as part of the
    variable name. Moisture (water) (WATER) (g), being the Tagname "WATERg".

-   Similarly, in the Indian FCDB, 2017 (IN17) which was obtained from
    the sharing agreement with the FAO team, the Tagnames were available
    in the FCDB and complemented with the FCDB documentation [REF]. Note
    that a great number of organic acids were reported (Table 9, from
    column 123), of which very few Tagnames were available.

-   In the Japan FCDB, 2015, Tagnames were also self-contained in
    original FCT file. However, manual renaming of some components,
    as it was done for the IN17, was needed. For instance, Protein,
    calculated from reference nitrogen_g to "PROCNTg", Lipid_g to
    "FATg", and Carbohydrate, total, calculated by difference_g to
    "CHOAVLDFg".
    
-   In Kenya FCDB, 2018, Tagnames were also self-contained in the original database and also have been previously reviewed for the MAPS project. Only small fixes were needed to be adapted for this project (i.e., adding the units along the Tagnames or filtering columns that were mostly empty (FA)).

-   In Western Africa FCDB, 2019, Tagnames were provided and, as the
    KE18, have been also previously reviewed and pre-processed for MAPS.
    Note that the edible portion is divided into two variables: EDIBLE1
    and EDIBLE2, we are using EDIBLE1 as the `Edible_factor_in_FCT`
    variable following FAO Fisheries Global NCT, 2022.

-   In Brazil FCDB, 2013 Tagnames were provided by the FAO team as it was
    previously processed. Hence, only renaming of the information variables (e.g., food id = fdc_id) was needed.

-   In uFiSh FCDB, 2016, it included Taganames and we only needed
    to allocate the edible portion to Edible coefficient, from whole to fillet/flesh (EDIBLE1), following the FAO Fisheries Global NCT, 2022 and as it was done in WA18.

For others datasets included, Tagnames were not reported and they needed to be
generated using the information reported in the Fisheries Global NCT, 2022, and the documentation accompanying each FCT.

-   In the Denmark FCDB, 2019 documentation [REF](<https://frida.fooddata.dk/pdf/en-frida-2-introduction_20-12-2016-esax.pdf>), we found that the variable named as "vitamin A (RE)" was in reality the
Vitamin A RAE, and hence the corresponding Tagname was VITA_RAEmg,
similarly vitamin D was VITDEQ, according to Denmark FCDB, 2019
documentation and FAO/INFOODS, 2012. 

-  In New Zealand FCDB, 2018, renaming was manually done, additionally information about vitamin D was uncertain.

-  In the UK FCDB, Tagnames were added manually. In addition, some things
that were noted were: Alcohol was reported as volumen (Alcohol (ALC) g/
mL (w/v)). We did not transformed the units because: 1) alcohol was not included as one of the NVs to be reviewed and, 2) we finally did not include this FCDB into the final Fishery Global NCT, 2022. However, for the future, if we want to change to match the 100g EP values, we could re-calculate ALC variable by using Eq.1 and the density of each food which is provided in the FCDB (specific_gravity). Vitamin D was renamed to the VITDmg Tagname, however re-calculation using D3 and 25-hydroxy D3 values would be recommended when
possible. Finally, for the DHA and EPA, we used cis- version, but it should
be reviewed. This was based on the documentation available in the
Composition of Foods integrated dataset: PHE (Public Health England)
(2021) Composition of foods integrated dataset (CoFID). Available at:
<https://www.gov.uk/government/publications/composition-of-foods-integrated-dataset-cofid>
(accessed 28/03/2022).

-   In the US FCDB, renaming was also done manually and based on the documentation and the information provided in the Fisheries Global NCT, 2022. We identified the Tagnames in one of the attached file, hence for the future a comparison/ update using those Tagname could be done. 

-   In Bangladesh FCDB, 2013 we renamed the variables to the Tagnames
manually, VITE definition was not clear and hence, in future should be checked.

-   In Norwegian FCDBs, 2021, variables were manually renamed to the Tagnames.
This was included for this project hence a more in-depth review of each
component was needed and it is described in the **Section X.X of the NO21 documentation**

##### Removing parentheses and brackets 

In JA15 parentheses were used to denote "estimated values". According to JA15 documentation, those were used to "Composition values estimated based on values listed in the food composition tables of foreign countries, raw material blending ratio recipe, etc. are indicated by placing them in parentheses. For some of Cereals, Fruits and Mushrooms, composition values derived from estimation or calculation using the listed values of similar foods are indicated by placing them in parentheses." (REF).

In KE18, brackets (\[\]) denoting not preferred values, as for the KE18 reporting: "For alternative analytical method or lower quality. Data from non-preferred tagnames were presented only where preferred tagname data was not available." (REF).

Bracket for low quality values were removed or generated a new column
with the alternative analytical method for UF16, WA19 and KE18. In addition, for the later, a star ("\*") was used to identify items fortified with folic acid. Note some issues on the reporting of this fortification were found previously, however the data was not reviewed in-depth because folate and derivates were outside the scope of this work. When we removed the brackets, for future development accounting for the impact of the "estimated values" should be studied. 

#### Transforming values

##### Characters in nutrient values
**NOTE: Maybe good to convert into table**

Trace and/or below detection limit were reported and converted to zero in the following FCTs: IN17 ("<LOD"), BA13, BR11, JA15 ("Tr"), and KE18, UK16, and WA19 ("tr"). 

Characters were used to report missing values were "nv" in DK19, "-" in JA15, and "M" in NO21. All were transformed to `NA`. 


##### Unit conversions and re-calculation

In addition, some components were reported in different units as those
reported in the Tagnames, hence unit conversion operations were performed. For
instance:

In the Indian FCDB, 2017, fatty acids needed to be converted from mg/100g EP to g/ 100g EP, and amino acids from g/100g PROT to mg/100g EP (Eq.2.3).

Similarly, in the Japan FCDB, 2015, we need to change the unit of the fatty acids from mg to g, and the edible portion was calculated from refuse (Eq.3.1).

Moreover, when (substantial) changes, like those mentioned above, were
made to the values reported in the original FCDBs, documentation was
done on GitHub so future users could review our decisions:

Documentation in GitHub [Issue #9](https://github.com/LuciaSegovia/UoN-FAO/issues/9), [Issue #10](https://github.com/LuciaSegovia/UoN-FAO/issues/10). 

#### Combined variables (standardised):

This was done after merging all the dataset together, and it's reported on the `QC.R`. 

Here's the list of the standardised variables as reported in the Fisheries Global NCT, 2022, of which we only reported those in bold (e.g., VITB6_mg_standardised, FAT_g_standardised)

 - SOP_own = Sum of Proximate
 - ENERCkJ_standardised
 - ENERCkcal_standardised
 - CHOAVLDFg_standardised
 **- FAT_g_standardised**
 - FIBTGg_standardised
 - VITAmcg_standardised
 - VITA_RAEmcg_standardised
 - CARTBEQmcg_standardised
 - VITEmg_standardised
 - THIAmg_standardised
 - FOLDFEmcg_standardised
 **- VITB6_mg_standardised**
 **- NIAEQmg_std**
 - NIAEQmg_standadised
 - NIATRIPmg_standardised
 - VITDmg_standardised

Scripts with all renaming and other transformation the steps for each FCDBs 

#### Updating the Global Fish NCT:

##### 1.  Merging all the data together

Once the FCTs/FCDBs were loaded and harmonised, we combined them into
a single data file. The UK21 was excluded from the Fishery Global NCT, 2022 hence ICS FAOSTAT code were not available for the fish in that FCT. For identifying fishery and fishery products in that FCT we used the `food_group`: White fish (JA), Fatty fish (JC), Crustacea  (JK), Molluscs (JM), and Fish products and dishes (JR). The id could be used as well, as all fisheries `fcd_id` started with "16". 

##### 2.  Allocating the FAO codes: ISSCAAP, alpha-three, and ICS FAOSTAT.

Fish matching between FCTs fish entries and ICS FAOSTAT, alpha-three code and ISSCAAP code was previously done in the Fishery Global NCT, 2022. Hence, we used that information to generate a dataset ("ics-code_fish-code.csv") with information on: `ICS.FAOSTAT.SUA.Current.Code`, `source_fct`, `NDB_number`, `fdc_id`, `Food.description`, `Scientific.name`, `ISSCAAP.Group`, and `X3.alpha.code`. Minor adjustment were needed, for instance, the `fcd_id` in the Fishery Global NCT, 2022 were as in the "original" KE18. In order to match our `fdc_id` we need to remove the first "0" in each code. For the US19, we needed to change the `fdc_id` to `NDB_number` for the merging. Then, we merged the dataset derived from Fishery Global NCT, 2022 with information on each fish entry ISSCAAP, ICS FAOSTAT and alpha-three codes (`ics_code_file`) and the newly Fisheries nutrient table (`fish_fct`) by each fish entry id. (`fdc_id`). That operation resulted in a dataset with all the fishery products and its corresponding ICS FAOSTAT code, alpha-three code and ISSCAAP code.

After merging, we kept only fish that were allocated with ICS code, hence, around 318 fish entries were filtered out, mainly from UK21 but some from NO21 (see NO21_docu.R) as well:

`fish_fct %>% filter(is.na(ICS.FAOSTAT.SUA.Current.Code)) %>% count(source_fct)`

```{r}

source("merging_all.R")

```

##### 3.  Calculating the component values per ICS FAOSTAT fish category

After merging and filtering the fishes, and before calculating the nutrient values of each ICS FAOSTAT fish category, we performed some visualisation (e.g., data availability) of the data and some quality check for most of the nutrient values. Those quality check are reported in `visualisation.R` and `QC.R`.


###### Proteins



###### Carbohydrates

-   Five FCTs/FCDBs did not report CHOAVLDFg
    (`fao_fish_fct %>% filter(is.na(CHOAVLDFg)) %>% distinct(source_fct)`).
-   Of those, two reported CHOCDFg instead
    (`fao_fish_fct %>% filter(is.na(CHOAVLDFg), !is.na(CHOCDFg))`)
-   Other two reported CHOAVLg
    (`fao_fish_fct %>% filter(is.na(CHOAVLDFg), !is.na(CHOAVLg)) %>% count(source_fct)`).
-   IN17 did not reported content in CHO and fibre for fish and
    products. Although, it could be calculated as CHOAVLDFg, or we could
    also assumed to be zero.
-   CHOAVLMg and CHOCSMg could be removed from the final database as it
    is of no use and/or relevance for fishes.

###### Fat

-   Eight FCTs/FCDBs did not report FATg
-   Three reported FAT_g
    (`fao_fish_fct %>% filter(is.na(FATg), !is.na(FAT_g)) %>% count(source_fct)`)
    while four reported FATCEg
    (`fao_fish_fct %>% filter(is.na(FATg), !is.na(FAT_g)) %>% count(source_fct)`)
-   WA19 reported mainly FATg but 57 entries were reported as FATCEg
    (21%)

- We combined those Tagnames into `FAT_g_standardised` as reported in section 4.4. 


###### Fibre

-   Fibre, crude (FIBCg) could be removed, it did not added information on any missing (FBGTg). For the future development, check whether NA values for fibre could be assume equal to zero (see Next steps: Assumed zero section).


###### Alcohol

-   Ten FCTs/FCDBs did not report ALCg, of the three that did report it,
    the mean value was zero. For the future, we can hence assume alcohol = zero for fish and products.


Of the nutrients used to calculate energy, the most important for fish and fish products are proteins, fat, water and ash. Hence, the review of those components should be carried more in-depth than the other, as for CHO, fibre, and alcohol could be assumed zero, for most of fish and shellfish products, with the exceptions of preparations, fermented or molluscs (FAO/INFOODS, 2012). 

###### 1) Niacin

Niacin (`NIAmg`) was reported in all FCDBs but in WA19, in which niacin equivalent (`NIAEQmg`) was reported. Because tryptophan was reported in the WA18 `NIAmg` could be re-calculated to `NIAmg_std`. 

We check for "extreme values", and after plotting the values comparing the values, overall all values seemed plausible. 

Note that for "Other pelagic fish, cured", one of the high values is having a great influence in the final mean value. For instance, if we remove values that higher than 40, the mean of that category is 15mg/100g instead of 30mg/100g. The rest of the categories remained more or less the same. This could be solved by including more entries (more values) for that fish category, as currently there are only two entries for that category. 

###### 2) Copper

Copper was reported by all FCDBs but KE18. Overall it had a very high data availability with very few missing values.

Octopus from IN17 (fdc_id = R003) looks very high compared with other octopus values. Double-check the value. 

Oyster from UF16 (fdc_id = 093013) also seems high compared with other oyster but seems plausible given the type of food. Double-check the value.

In addition, there is one item that is driving the very high mean for the category "Cephalopods, cured", again, more values should be sought for that category.


###### 3) Vitamin B6

Vitamin B6 was reported by all the FCDBs but KE18. Most of the FCDBs
used analytical values (VITB6Amg), followed by calculated (VITB6Cmg) and
only the Norwegian FCDBs used a mixed of values (VITB6_mg).

Then, as reported in section 4.4., VITB6Cmg, VITB6Amg, and VITB6C_mg were combined into `VITB6C_mg_standardised`. 

###### 4) Vitamin B12

Vitamin B12 was reported in nine FCTs. JA15 seemed to report higher values than the other FCTs, including most of the high-end values. 

We found that "Molluscs, excluding cephalopods, frozen", "Molluscs, excluding cephalopods, fresh", and "Molluscs, excluding cephalopods, canned" had the highest Vit. B12 values. We found that clams from JA15 had the highest values for the "Molluscs, excluding cephalopods, frozen". However, when we plotted the data it seemed to have a normal distribution, and hence we are not considering them as outliers.  

Then, we realised that most of the values for that category were missing values. More values are needed as similar (and even higher) concentration of Vit.B12 have been previously reported in literature(Watanabe et al., 2014, 10.3390/nu6051861). Since, the water content is very low as well the value seemed plausible.

The category "Cephalopods, preparations nei" have very high variability, it seemed that octopus had higher Vit.B12 than other cephalopods in the category, also most of the high values were sourced from UF16, while the highest was from US19. It maybe that the UF16 values were imputed from US19. 

###### FAT

In order to compare the DHA and EPA with the fat content, we needed to combine several fat Tagnames.

###### 5) DHA 

Of the seven FCTs that reported DHA, some "outliers" could be seen in the US19 and DK19 with values higher than 5g of DHA. 

When we checked for outliers, all the very high values (>4g) were reported in oils, hence they seemed plausible values. For the other values higher than the Q4, most corresponded to fish and preparation with high fat content (>5%) and contained relatively low water content (<70%). 

Apparently, it did not seem to have major differences when we compared the mean values of DHA and Fat with and without values completeness. 

However when we checked the percentage of difference between fat content of the items with DHA values and without, some concerning differences were shown. This was particularly true for very heterogeneous SUA categories were also limited number of items were included, for instance, "Aquatic mammals, meat", "Aquatic animals nei, fresh". We could not find values for DHA for those items in literature or other FCTs, hence this remainded as an area were improvements are needed. 

"Aquatic plants, dried", there are many `NA`, however, for those that values are available DHA is zero. Maybe we could "assume zero" for those values. 


######6) EPA

Results of the EPA checks were very similar to those for DHA. 

###### 7) Selenium

#**CHECK BC VITB12 wasn't loading in JAPAN**

When we checked the values one value, "Tuna, raw" (04.043) from the NO21, seemed very high compared with other tuna values. There were other high values but they seemed to be still within plausible ranges. 

When we plotted the results by FCT, and by fish type we could see that NO21 reported overall the highest Se median value with "extreme values", coming consitently from that FCT, followed by NZ18, and US19. 

Some values to be checked from NZ18 was "Lobster, cooked" (T39), and from the NO21 ("04.043", "04.333", "04.107", "04.055")). 

When we checked by SUA category, we identified the "Aquatic mammals, preparations nei" having the highest mean value. It only reported information for five entries of which only two had a `SEmcg` value. We could imputed (35055) with (35164), by water-adjusting and assuming the oil has no contribution to `Semcg`. Although, due to the low water content and overall high content in Se of the category, it would only increase the mean. Again from the "Aquatic mammals, meat", we only have 3 values out of 15 values, which makes the mean quite high, although we can assess if that mean is high or is in line with the overall category mean due to the lack of data availability. 



#Data avaibility per each ICS FAOSTAT fish category

We identified missing values at ICS FAOSTAT fish category, only for Se in four ICS FAOSTAT fish categories: 

`print(missing)`

For **Cephalopods, cured**, we used the water-adjusted Se mean concentration of the following items, using the WATERg and SEmcg of: (All from JA15)

|fcd_id| food_desc| WATERg|SEmcg|
|-|-|-|-|
|10342| Mollusks, neon flying squid, raw | 79.3 |28|
|10345| Mollusks, Japanese common squid, raw [Syn. short-finned squid] |
80.2| 41|

To "complete" the Se concentration of the following "Cephalopods, cured":

[1] "Mollusks, firefly squid, seasoned and smoked"\
[2] "Mollusks, processed squid, "Surume" dried squid"\
[3] "Mollusks, processed squid, "Saki-ika" dried, seasoned and shredded
squid"
[4] "Mollusks, processed squid, seasoned and smoked"

For **Cephalopods, canned**, there was only one item under that category: *"Mollusks, Escargot Apple snails, canned in brine"*. However that snails should be recorded under "Molluscs, excluding cephalopods", according to the description in the documentation (REF). Hence, to populate this category, we decided to use the fish entries from "preparations nei", which were described as "boiled/ moist heat" (n=22). This was decided based on:

a) The description of *Cephalopods, Canned* in the FAO documentation, which is "Raw or precooked meat from gutted or ungutted animals, canned in brine, oil or other medium (all edible) treated at temperatures adequate to ensure sterilization. Examples are: vacuum-packed small octopus, canned squid in own ink or oil.".
b) Nutrient values reported on the Spanish FCTs for "squids, fresh" and "squids, canned" were very similar. Unfortunately, there was no information for Se,
c) And, same fish should be a better representation, despite the cooking preparation, than a different food, particularly when they are split into different categories. In addition, when perfect match is not possible the mean of several foods is recommended (FAO/INFOODS, 2012)

NOTE: Some concerns related to the edible portion, so in the "canned" description we are assuming that it is all edible, and hence 100g of a canned squid would have X% of squid and Y% of the medium. This is not captured by the boiled, or fresh product.

For **Aquatic animals, nei, cured and preparation nei**, when we checked the items included in that category there were mostly Sea urchin, jellyfish, sea cucumber and sea squirt from the Japan FCT, 2019. However, when we tried to identify Se values from the "Aquatic animals, nei, fresh" category, we realised that it was very diverse category, including from crocodiles to sea cucumber. The only Se information that was available and comparable was for sea cucumber,
thus we used `WATER` and `SEmcg` to water-adjust the Se concentration of: "Sea cucumber, raw" (WATERg = 92.2, SEmcg = 37) for imputing `SEmcg` in: 

[1] "Sea cucumber, "Konowata" salted and fermented viscera"
[2] "Sea urchin, "Tsubu-uni" salted whole gonads"\
[3] "Sea urchin, "Neri-uni" salted whole gonad paste"\
[4] "Jellyfish, Salted, desalted"
[5] "Sea squirt, "Shiokara" salted and fermented meat and viscera \\[Syn. Ascidian\\]"

The median Se conc. of the category is 14.1mcg.

For **Crustaceans, cured**, we needed to impute Se concentration for "Crustacean, Sakura shrimp, dried", for which we used (10415) "Crustacean, whiteleg shrimp, raw" (WATER = 78.60, SEmcg = 27). And, for "Shrimp (crayfish), whole, dried",  we calculated mean/median of UF16 values as it was reported in the references (UF1(092004)), plus the following: "European/Euro-American/Oceanian crayfishes, flesh, raw (n.s.)", "European/Euro-American/Oceanian crayfishes, wild, flesh, raw (n.s.)", "European/Euro-American/Oceanian crayfishes, farmed flesh, raw". According to Vásquez-Caicedo et al. (2008) and Bognar (2002), retention factors for Se are between 0.9-1, with dry heat (braise, fry in oven, broiling, grilling) being 1.

Information regarding the values used to impute/calculate the `SEmcg` for those items can be found un the column `commet`. 

#Questions:

Why using PROCNT(g) instead of PROTCNTg?

Snail (Helix pomatia) - under Molluscs, excluding cephalopods (that's
ok) and under Cephalopods (why?)

### Final Quality Checks

Because this FCT is an extension of a existing FCT, we need to perform some extra-control steps to ensure that values are a good representation of the data shown. We also need to perform those extra steps as we have added some new fish entries, and hence, we need to make sure that reconciliation with all the other nutrients are performed. 

#### Water content

#### DHA and EPA concetration

We calculated the mean and other summary stats for the each fish category in total, and only for the entries that reported the DHA and EPA. This was to evaluate the impact of missing values into the final results. 

We also calculate the percentage of FA (i.e., the corresponding percentage of each FA to the total FAT). To do so we used the following equation. Note that we used the `FAT_g_standardised` variable which is a combination of the Tagnames (and different extractions): FAT, FAT_, and FATCE. In addition, FATCE is not recommended for use in unit conversion (FAO/INFOODS, 2012). 

Eq.: FA(%) = FAg/FATg x 100 , (F22D6N3g_perc = F22D6N3g/FAT_g_standardised*100)

However, as it is not going to be used to report values, and only for checking, we proceed to compare the percentage of each FA for the whole dataset and only those entries with FA values. 

We added information on the difference in percentage for each fish category, and we found that
for DHA only 19 fish categories (out of 95) had a difference smaller than zero (0), that ranged from -0.55 to -0.17, and no difference value was higher than zero. Whereas for EPA, we found five values smaller than zero (-0.02 to -0.11), and two values larger than zero: 1.13, and 1.58 for "Aquatic plants, dried", and "Aquatic plants, preparations nei". For which small number of entries were included, and a considerably diverse fishery category. 

We also calculate the percentage of relative coverage (number of entries reporting FAT_g_standardised/ number of entries reporting the DHA/EPA *100). Thirty-one and thirty-two fish categories had lower than 60% coverage for DHA and EPA respectively. 




## Next steps: Functions development and improvemnts

#### Renaming variables

Overall
- Allocate CDNO ontology codes to NV.


B13:
- Get CARTBEQ or [CARTB] into two columns.
- Remove function and change to pre-loading functions from source(functions.R).

DK19
-Rename all other NV w/ Tagnames.
-Find a better renaming way.

NZ18
-Rename all NV w/ Tagnames, including sugars and Fatty Acids. 

US21
- Using data embeded in US21 to allocate the Tagnames for all NVs. 

#### Characters in nutrient values

-   A function that generate different columns for different analytical
    method of the same/similar component. E.g., Extracting alternative
    analitycal method (i.e., NIA for Niacin) that are provided in the
    same column of the preferred analytical method (i.e., NIAEQ for
    Niacin) but within bracket.

Observed in BA13, KE18, WA19,

-   A function that extract the "estimated" values and/or low quality
    values between brackets or parenthesis, and it generate a metadata
    info on that. Eg. A new column with a string providing information
    on which values were low quality and/or estimated

JA11: There is a use of bracket to denote "estimated values". We are
removing the brackets but we should look at how to keep a record and
what's the meaning of estimated values.

KE18: "Low quality" values were reported within brackets.

-   A function to change Trace values, expressed as "tr", "Tr", "\<LOD",
    to 0 (zero).

UK21 - Check "N" values (found in Edible_portion)
    
#### Sum of proximate (SOP)

Sum of proximates (=Σof water + protein + fat + available
carbohydrates + dietary fibre + alcohol + ash)

Ranges: Preferable: 97 - 103 g (Greenfield and Southgate, 2003,
FAO/INFOODS, 2012)) Acceptable: 95 - 105 g (FAO/INFOODS, 2012))

#### Nutrient values

##### DHA and EPA

Fats and fatty acids concentration in fish and aquatic animals varies widely. These components are very sensitive to many environmental, seasonal and other changes. Hence, imputing missing values is not optimal. Therefore, including more fishery entries with minimal missing values should be sought in the future. 

##### Energy

We will generate a function that will calculate energy in kcal and
energy in kj using the FAO conversion factors and the fao tagnames that
will be used to generate the columns: `ENERCkJ_standardized` and
`ENERCkcal_standardized` for the FAO fishery products database and for
MAPS to calculate energy used.

#### Carbohydrates available by difference (CHOAVLDF(g)\_standardized)

100 - (weight in grams [water + protein + fat + ash + alcohol + dietary
fibre] in 100 g of food)

#### Niacin Eq.

Total niacin equivalent (NIAEQ) = niacin preformed (NIA) + 1/60
tryptophan (TRP)

#### Unit conversion

Unit conversion (when possible) to match Tagnames (per 100g EP FW)

- Alcohol(w/v) to (w/100g) = ALC (g/100mL) / density (g/mL) = ALC (g/100 EP)



###### Dealing with missing nutrient values

##### Assumed zero

-Fibre: We could check whether they are listed under fish or  fish preparations. If preparations then check what type of preparation, otherwise we could assume zero.

Current mean of FIBTGg =
`fao_fish_fct %>% filter(!is.na(FIBTGg)) %>% pull(FIBTGg) %>% as.numeric() %>% mean(., na.rm = T)`.

- Alcohol: We could assume zero.

##### Imputing nutrient values
